### 2023-09-18 manuscripts/xxx-2024/main.tex
start working on second paper

\documentclass[conference]{IEEEtran}

\usepackage[colorlinks,allcolors=blue]{hyperref}
\usepackage{amsfonts}
\usepackage{graphicx}

\begin{document}

    \title{A formal theory of digital twins}

    \author{
        \IEEEauthorblockN{Georg Hackenberg} 
        \IEEEauthorblockA{School of Engineering\\University of Applied Sciences Upper Austria\\4600 Wels, Austria} 
        \and 
        \IEEEauthorblockN{Alican Tüzün} 
        \IEEEauthorblockA{Josef Ressel Center for Data-Driven Innovation in Business Modeling\\University of Applied Sciences Upper Austria\\4xxx Steyr, Austria}
    }

    \maketitle

    \begin{abstract}
        TODO
    \end{abstract}

    \section{Introduction}
    \label{section:introduction}
    TODO~\cite{asf}

    \subsection{Research objective}
    More systematic development of high-quality digital twins in different application domains.

    \subsection{Research question}
    Unambiguous definition of the term "digital twin"?

    \subsection{Research methodology}
    Literature review + formalization

    \section{Related work}
    TODO

    \section{A digital twin theory}
    \label{section:theory}
    TODO

    \subsection{Basic concepts}

    \subsubsection{Time}
    \[
        T
    \]

    \subsubsection{State}
    \[
        S    
    \]

    \subsubsection{Stream}
    \[
        \overrightarrow{S}: T \rightarrow S
    \]

    \subsection{Our model of the real world}

    \subsubsection{State set of the environment}
    \[
        S_{Env}    
    \]

    \subsubsection{State set of the product}
    \[
        S_{Pro}    
    \]

    \subsubsection{Behavior function of the environment}
    \[
        B_{Env}: \overrightarrow{S_{Pro}} \rightarrow \overrightarrow{S_{Env}}
    \]

    \subsubsection{Behavior function of the product}
    \[
        B_{Pro}: \overrightarrow{S_{Env}} \rightarrow \overrightarrow{S_{Pro}}
    \]

    \subsubsection{Actual behavior}
    tuple of product state stream and environment state stream
    \[
        (\overrightarrow{S_{Pro}}, \overrightarrow{S_{Env}})
    \]
    such that state streams are consistent, i.e.
    \[
        B_{Env}(\overrightarrow{S_{Pro}}) = \overrightarrow{S_{Env}}
    \]
    and
    \[
        B_{Pro}(\overrightarrow{S_{Env}}) = \overrightarrow{S_{Pro}}
    \]

    \subsection{Digital twin}
    TODO

    \section{Conclusion}
    \label{section:conclusion}
    TODO

    \bibliography{main}
    \bibliographystyle{plain}

\end{document}